---
title: "ChallengeV_Answers"
author: "Jacobo Dobon"
date: "2026-03-05"
output:
  
  md_document:
    variant: gfm
  html_document:
    toc_float: true
---

## [Link](https://github.com/Jake-Ramirez/Data_wrangling)

### 1. Download two .csv files from Canvas called DiversityData.csv and Metadata.csv, and read them into R using relative file paths.

```{r}
library(tidyverse)

Diversity_Data <- read.csv("DiversityData.csv")
Metadata <- read.csv("Metadata.csv")
```


### 2. Join the two dataframes together by the common column ‘Code’. Name the resulting dataframe alpha.

```{r}
alpha <- left_join(Diversity_Data, Metadata, by = "Code")
head(alpha)
```


### 3. Calculate Pielou’s evenness index: Pielou’s evenness is an ecological parameter calculated by the Shannon diversity index (column Shannon) divided by the log of the richness column. 

  a.	Using mutate, create a new column to calculate Pielou’s evenness index.
  b.	Name the resulting dataframe alpha_even.
```{r}
alpha_even <- alpha %>%
  group_by(shannon, richness) %>%
  mutate(pielou = shannon/log(richness))
head(alpha_even)
```


### 4. Using tidyverse language of functions and the pipe, use the summarise function and tell me the mean and standard error evenness grouped by crop over time.

```{r}
alpha_average <- alpha_even %>%
  group_by(Crop, Time_Point) %>%
  summarise(Mean.pielou = mean(pielou), # calculating the mean richness, count, stdeviation, and standard error
            n = n(), 
            sd.dev = sd(pielou)) %>%
  mutate(std.err = sd.dev/sqrt(n))
head(alpha_average)
```


### 5. Calculate the difference between the soybean column, the soil column, and the difference between the cotton column and the soil column

```{r}
alpha_average2 <- alpha_average %>%
  select(Time_Point, Crop, Mean.pielou) %>%
  pivot_wider(names_from = Crop, values_from = Mean.pielou) %>%
  mutate(diff.cotton.pielou = Soil-Cotton, diff.soybean.pielou = Soil-Soybean)
head(alpha_average2)
```


### 6. Connecting it to plots

```{r}
plot <- alpha_average2 %>%
  select(Time_Point, diff.cotton.pielou, diff.soybean.pielou) %>%
  pivot_longer(c(diff.cotton.pielou, diff.soybean.pielou), names_to = "diff") %>%
  ggplot(aes(x=Time_Point, y=value, color = diff)) +
  geom_line() +
  theme_minimal() +
  xlab("Time (hrs)") +
  ylab("Difference from soil in Pielou's evenness")
print(plot)

```







