---
title: "CodingNotesWrangling"
author: "Jacobo"
date: "2026-03-13"
output: html_document
---
Loading tidyverse since I already have it installed
```{r}
library(tidyverse)

```

We read the .csv file, I modified the path to match my folders. We use str()to explore the object.
```{r}
microbiome.fungi <- read.csv("Bull_richness.csv")
str(microbiome.fungi)
```

Here we use the select funtion to only work with certain columns. We use ":" when for specifying a desired range of columns we want to select in the order they already are.
```{r}
microbiome.fungi2 <- select(microbiome.fungi, SampleID, Crop, Compartment:Fungicide, richness)
str(microbiome.fungi2)
```

Now we use filtefr() which is another way of subsetting
```{r}
head(filter(microbiome.fungi2, Treatment == "Conv."))
```

The good thing about filter is that we can use conditional to selects rows with certain values in different columns
```{r}
head(filter(microbiome.fungi2, Treatment == "Conv." & Fungicide == "C"))
```

and another example
```{r}
head(filter(microbiome.fungi2, Sample == "A" | Sample == "B")) # samples A or B
```

Now we are gonna use mutate() to create new columns
```{r}
microbiome.fungi2$logRich <- log(microbiome.fungi2$richness) 
# Create a new column called logRich
head(mutate(microbiome.fungi2, logRich = log(richness)))
```

And if we use paste we can create a new column that integrates the data in other two
```{r}
head(mutate(microbiome.fungi2, Crop_Treatment = paste(Crop, Treatment)))
```


In R the pipe sign is %>%, it used to use the output of a function directly as input of the next one. Here I directly add the selecting and group by part needed for calculating stats as well as creating the column for the results and plotting.
```{r}
plot1 <- microbiome.fungi %>%
  select(SampleID, Crop, Compartment:Fungicide, richness) %>% # selecting columns
  group_by(Treatment, Fungicide) %>% # grouping by treatment and fungicide to later calculate summary stats by group
  mutate(logRich = log(richness)) %>% # creating a new column of the log richness
  summarise(Mean.rich = mean(logRich), # calculating the mean richness, stdeviation, and standard error
            n = n(), 
            sd.dev = sd(logRich)) %>%
  mutate(std.err = sd.dev/sqrt(n)) %>%
  ggplot(aes(x = Fungicide, y = Mean.rich)) + # adding in a ggplot
  geom_bar(stat="identity") +
  geom_errorbar( aes(x=Fungicide, ymin=Mean.rich-std.err, ymax=Mean.rich+std.err), width=0.4) +
  theme_minimal() +
  xlab("") +
  ylab("Log Richness") +
  facet_wrap(~Treatment)
```

Now lets select sample Id and richness on one side and sample ID again and all of the other columns that are not richness and use left_join to add richness from the only column in common
```{r}
# selecting just the richness and sample ID
richness <- microbiome.fungi %>%
  select(SampleID, richness)

# selecting columns that don't include the richness
metadata <- microbiome.fungi %>% 
  select(SampleID, Fungicide, Crop, Compartment, GrowthStage, Treatment, Rep, Sample)

head(left_join(metadata, richness, by = "SampleID"))
```

Pivoting
```{r}
microbiome.fungi %>%
  select(SampleID, Crop, Compartment:Fungicide, richness) %>% # selecting columns  filter(Class == "Sordariomycetes") %>%
  group_by(Treatment, Fungicide) %>% # grouping by treatment and fungicide to later calculate summary stats by group
  summarise(Mean = mean(richness)) %>% # calculates the mean per Treatment and Fungicide 
  pivot_wider(names_from = Fungicide, values_from = Mean) %>% # pivot to wide format
  mutate(diff.fungicide = C - F) # calculate the difference between the means. 
```

```{r}
microbiome.fungi %>%
  select(SampleID, Crop, Compartment:Fungicide, richness) %>% # selecting columns  filter(Class == "Sordariomycetes") %>%
  group_by(Treatment, Fungicide) %>% # grouping by treatment and fungicide to later calculate summary stats by group
  summarise(Mean = mean(richness)) %>% # calculates the mean per Treatment and Fungicide 
  pivot_wider(names_from = Fungicide, values_from = Mean) %>% # pivot to wide format
  mutate(diff.fungicide = C - F) %>%  # calculate the difference between the means. 
  ggplot(aes(x = Treatment, y = diff.fungicide)) + # Plot it 
  geom_col() +
  theme_minimal() +
  xlab("") +
  ylab("Difference in average species richness")
```




